---
title: "ESCA drug associations for perturbation-response genes"
author: "Michael Schubert (schubert@ebi.ac.uk)"
---

```{r code, include=FALSE}
library(dplyr)
b = import('base')
io = import('io')
ar = import('array')
st = import('stats')
df = import('data_frame')
plt = import('plot')
gdsc = import('data/gdsc')
util = import('./util')

INFILE = commandArgs(TRUE)[1] %or% "../../scores/gdsc/speed_matrix.RData"
OUTFILE = commandArgs(TRUE)[2] %or% "ESCA_assocs.pdf"

tissues = gdsc$tissues()
Ys = gdsc$drug_response('IC50s')
ar$intersect(tissues, Ys, along=1)

esca = gdsc$tissues("ESCA")
scores = io$load(INFILE)
Yf = gdsc$drug_response('IC50s', min_tissue_measured=2, median_top=10, drop=TRUE)
ar$intersect(scores, esca, Yf, along=1)

# tissues as subsets
assocs.tissue = st$lm(Yf ~ scores) %>%
    filter(term == "scores") %>%
    select(-term) %>%
    mutate(adj.p = p.adjust(p.value, method="fdr"))

#fh535 = filter(assocs.tissue, Yf=="FH535" & scores=="VEGF")
```

```{r volcano, echo=FALSE}
assocs.tissue %>%
    mutate(label = paste(Yf, scores, sep=":")) %>%
    plt$color$p_effect(pvalue="adj.p", effect="estimate", dir=-1, thresh=0.2) %>%
    plt$volcano(p=0.2)
```

Volcano plot of associations between predicted pathway activities
and drug response. Effect size on the horizontal, FDR-adjusted
p-values on the vertical axis. Size of circles corresponds to the
number of cell lines screened with each drug.

The top hit suggests that expression of VEGF/PDGF-responsive genes
mediates resistance to FH535 (Tcf/beta catenin inhibitor) with
p=10^-5, FDR<1%.

<!--
Other hits with a higher FDR (~25%) but additional support include:

* PI3K confers resistance to MEK inhibitors (PD-0325901, RDEA1119,
  Trametinib). This association has additional support in e.g. BRCA.
* Hypoxia confers resistance to LAQ824 (HDAC inhibitor). This is
  also one of the most significant (FDR=1e-3) BRCA associations.
-->

```{r tissue_compare, echo=FALSE}
util$drug_range_box("Rapamycin", highlight="ESCA")
```

Box plot of the measured (range indicated in color) and inferred IC50s
of the FH535 drug in the different cell lines tissues screened.

```{r fits, echo=FALSE}
do_fit = function(tissue, drug, path, gene, label_top=5) {
    fits = df$assemble(
        drug = Yf[,drug],
        path = scores[,path],
        muts = gdsc$mutated_genes(intogen=TRUE, tissue=tissue) %>%
            ar$map(along=2, function(x) paste(names(x)[x], collapse=",")),
        gene = gdsc$basal_expression()[gene,]
    ) %>% na.omit()
    # <this should not be required once df$assemble works>
    fits$drug = as.numeric(fits$drug)
    fits$path = as.numeric(fits$path)
    fits$gene = as.numeric(fits$gene)
    # </>
    fits$label = gdsc$cosmic$id2name(rownames(fits))
    fits = mutate(fits, label = paste(label, muts, sep="\n"))
#    fits$label[fits$gene < b$maxN(fits$gene, 5)] = ""

    ggplot(fits, aes(x=path, y=drug, label=label, fill=gene)) +
        geom_smooth(method=stats::lm, se=FALSE, na.rm=TRUE, alpha=0.1) +
        geom_point(pch=21, size=5, colour="black", na.rm=TRUE) +
        scale_fill_gradient(low="white", high="black",
                            limits=c(min(fits$gene, na.rm=TRUE),
                                     max(fits$gene, na.rm=TRUE)),
                            guide = guide_legend(title = gene)) +
        geom_text(colour="#353535", size=2, vjust=-1.5, na.rm=TRUE) +
        theme_bw() +
        ggtitle(paste(tissue, ":", drug, "response and",
                      path, "activity")) +
        xlab(paste("predicted", path, "activity")) +
        ylab(paste("IC50 [log uM]"))
}

print(do_fit("ESCA", "Rapamycin", "EGFR", "EGFR"))
print(do_fit("ESCA", "Rapamycin", "MAPK", "MAP2K1"))
```

Linear association between predicted pathway activity and drug
response to FH535, with each point representing a cell line
coloured by PDGFR A/B expression.
