---
title: "MM drug associations for perturbation-response genes"
author: "Michael Schubert (schubert@ebi.ac.uk)"
---

```{r code, include=FALSE}
library(dplyr)
b = import('base')
io = import('io')
ar = import('array')
st = import('stats')
df = import('data_frame')
plt = import('plot')
gdsc = import('data/gdsc')
util = import('./util')

INFILE = commandArgs(TRUE)[1] %or% "../../scores/gdsc/speed_matrix.RData"
OUTFILE = commandArgs(TRUE)[2] %or% "MM_assocs.pdf"

scores = io$load(INFILE)
tissues = gdsc$tissues("MM")
Yf = gdsc$drug_response('IC50s', tissue="MM", stage=2, drop=TRUE)
ar$intersect(scores, tissues, Yf, along=1)

# tissues as subsets
assocs.tissue = st$lm(Yf ~ scores) %>%
    filter(term == "scores") %>%
    select(-term) %>%
    mutate(adj.p = p.adjust(p.value, method="fdr"))
```

```{r volcano, echo=FALSE}
assocs.tissue %>%
    mutate(label = paste(Yf, scores, sep=":")) %>%
    plt$color$p_effect(pvalue="adj.p", effect="estimate", dir=-1, thresh=0.2) %>%
    plt$volcano(p=0.2)
```

```{r range, echo=FALSE}
util$drug_range_box("Sorafenib", highlight="MM")
```

<!--
```{r fits2, echo=FALSE}
#TODO: volcano of diff trail genes + trail pathway exp, show is diff?
```
-->

```{r fits, echo=FALSE}
do_fit = function(tissue, drug, path, gene_, label_top=5) {
    fits = df$assemble(
        drug = Yf[,drug],
        path = scores[,path],
        muts = gdsc$mutated_genes(intogen=TRUE, tissue=tissue) %>%
            ar$map(along=2, function(x) paste(names(x)[x], collapse=",")),
#        gene = gdsc$basal_expression()[gene,]
        gene = io$load('../../scores/gdsc/gsea_reactome.RData')[,'Trail']
    ) %>% na.omit()
    # <this should not be required once df$assemble works>
    fits$drug = as.numeric(fits$drug)
    fits$path = as.numeric(fits$path)
    fits$gene = as.numeric(fits$gene)
    # </>
    fits$label = gdsc$cosmic$id2name(rownames(fits))
    fits = mutate(fits, label = paste(label, muts, sep="\n"))
#    fits$label[fits$gene < b$maxN(fits$gene, 5)] = ""

    ggplot(fits, aes(x=path, y=drug, label=label, fill=gene)) +
        geom_smooth(method=stats::lm, se=FALSE, na.rm=TRUE, alpha=0.1) +
        geom_point(pch=21, size=5, colour="black", na.rm=TRUE) +
        scale_fill_gradient(low="white", high="black",
                            limits=c(min(fits$gene, na.rm=TRUE),
                                     max(fits$gene, na.rm=TRUE)),
                            guide = guide_legend(title = gene_)) +
        geom_text(colour="#353535", size=2, vjust=-1.5, na.rm=TRUE) +
        theme_bw() +
        ggtitle(paste(tissue, ":", drug, "response and",
                      path, "activity")) +
        xlab(paste("predicted", path, "activity")) +
        ylab(paste("log IC50", drug))
}

print(do_fit("MM", "Sorafenib", "Trail", "Trail expression"))
```

Expression of Trail-induced genes (on horizontal axis) are
correlated with drug sensitivity in Multiple Myeloma (MM),
while expression of members of the apoptosis pathway (GSEA score
as shades of grey) is not.

Presence of TP53 mutations also does not correlate with either
drug response or Trail activity, so this suggests an independent
mechanism.

Taken together, we find a possible synergy between rTrail and Sorafenib
as treatment in MM.
