---
title: "SKCM drug associations for perturbation-response genes"
author: "Michael Schubert (schubert@ebi.ac.uk)"
---

```{r code, include=FALSE}
library(dplyr)
b = import('base')
io = import('io')
ar = import('array')
st = import('stats')
df = import('data_frame')
plt = import('plot')
gdsc = import('data/gdsc')
util = import('./util')

INFILE = commandArgs(TRUE)[1] %or% "../../scores/gdsc/speed_matrix.RData"
OUTFILE = commandArgs(TRUE)[2] %or% "SKCM_assocs.pdf"

tissues = gdsc$tissues()
Ys = gdsc$drug_response('IC50s')
ar$intersect(tissues, Ys, along=1)

skcm = gdsc$tissues("SKCM")
scores = io$load(INFILE)
Yf = gdsc$drug_response('IC50s', min_tissue_measured=2, median_top=10, drop=TRUE)
ar$intersect(scores, skcm, Yf, along=1)

# tissues as subsets
assocs.tissue = st$lm(Yf ~ scores) %>%
    filter(term == "scores") %>%
    select(-term) %>%
    mutate(adj.p = p.adjust(p.value, method="fdr"))

fh535 = filter(assocs.tissue, Yf=="FH535" & scores=="VEGF")
```

```{r volcano, echo=FALSE}
assocs.tissue %>%
    mutate(label = paste(Yf, scores, sep=":")) %>%
    plt$color$p_effect(pvalue="adj.p", effect="estimate", dir=-1, thresh=0.2) %>%
    plt$volcano(p=0.2)
```

Volcano plot of associations between predicted pathway activities
and drug response. Effect size on the horizontal, FDR-adjusted
p-values on the vertical axis. Size of circles corresponds to the
number of cell lines screened with each drug.

The top hit suggests that expression of VEGF/PDGF-responsive genes
mediates resistance to FH535 (Tcf/beta catenin inhibitor) with
p=10^-5, FDR<1%.

<!--
Other hits with a higher FDR (~25%) but additional support include:

* PI3K confers resistance to MEK inhibitors (PD-0325901, RDEA1119,
  Trametinib). This association has additional support in e.g. BRCA.
* Hypoxia confers resistance to LAQ824 (HDAC inhibitor). This is
  also one of the most significant (FDR=1e-3) BRCA associations.
-->

```{r tissue_compare, echo=FALSE}
util$drug_range_box("FH535", highlight="SKCM")
```

Box plot of the measured (range indicated in color) and inferred IC50s
of the FH535 drug in the different cell lines tissues screened.
Melanoma cells are, by median, the most sensitive tissue.

Here we show that melanomas are the most sensitive tissue for the
drug FH535. Combined with our resistance marker, this offers a way
of further stratifying cell lines (and in turn maybe patients) when
there is no mutation marker available.

```{r fits, echo=FALSE}
do_fit = function(tissue, drug, path, gene, label_top=5) {
gex = gdsc$basal_expression()
    fits = df$assemble(
        drug = Yf[,drug],
        path = scores[,path],
        muts = gdsc$mutated_genes(intogen=TRUE, tissue=tissue) %>%
            ar$map(along=2, function(x) paste(names(x)[x], collapse=",")),
        gene = log10(10^gex["PDGFRA",]+10^gex["PDGFRB",])
    ) %>% na.omit()
    fits$label = gdsc$cosmic$id2name(rownames(fits))
    fits$label[fits$gene < b$maxN(fits$gene, 5)] = ""

    ggplot(fits, aes(x=path, y=drug, label=label, fill=gene)) +
        geom_smooth(method=stats::lm, se=FALSE, na.rm=TRUE, alpha=0.1) +
        geom_point(pch=21, size=5, colour="black", na.rm=TRUE) +
        scale_fill_gradient(low="white", high="black",
                            limits=c(min(fits$gene, na.rm=TRUE),
                                     max(fits$gene, na.rm=TRUE)),
                            guide = guide_legend(title = gene)) +
        geom_text(colour="#353535", size=2, vjust=-1.5, na.rm=TRUE) +
        theme_bw() +
        ggtitle(paste(tissue, ":", drug, "response and",
                      path, "activity")) +
        xlab(paste("predicted", path, "activity")) +
        ylab(paste("IC50 [log uM]"))
}

print(do_fit("SKCM", "FH535", "VEGF", "PDGFRA+B"))
#print(do_fit("SKCM", "FH535", "VEGF", "PDGFRB"))
#print(do_fit("SKCM", "FH535", "VEGF", "PPARG"))
```

Linear association between predicted pathway activity and drug
response to FH535, with each point representing a cell line
coloured by PDGFR A/B expression.

Not only is higher predicted pathway activity significantly associated
with poorer drug response, but a subset of resistant cell lines have
ovexpressed their PDGFR receptor.

This suggests a receptor overexpression mediated induction of PDGF
signaling and thus drug resistance, which he hypothesize could be
abrogated by either knocking down the receptor or selectively
inhibiting it.
