---
title: "Pathway scores in Basal and Perturbed Condition"
author: "Michael Schubert (schubert@ebi.ac.uk)"
output:
  pdf_document:
      number_sections: yes
      toc: yes
---

# Methods

## Volcano Plots

... write about methods here ...

## Individual Fits

... write about methods here ...


```{r setup, include=FALSE}
library(dplyr)
b = import('base')
io = import('io')
ar = import('array')
st = import('stats')
plt = import('plot')
util = import('./util_1')

overview = import('../analyses/speed_overview/plot_bars')
cluster = import('../analyses/speed_raw/plot')
roc = import('../analyses/speed_raw/roc')
stability = import('../analyses/speed_bootstrap/stability_plot')
```

# Size of the Data Set

```{r dataset_size, echo=FALSE}
overview$do_plot()
```

# Global Structure of Perturbation Experiments

```{r global_structure, echo=FALSE}
cluster$do_plot()
```

# Pathway scores upon Perturbation

## Signaling Footprints

```{r perturb_speed_matrix, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('speed_matrix')
```

## Gene Ontology

```{r perturb_go, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('gsea_go')
```

## Reactome

```{r perturb_reactome, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('gsea_reactome')
```

## BioCarta

```{r perturb_biocarta, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('gsea_biocarta')
```

## SPIA

```{r perturb_spia, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('spia')
```

## Pathifier

```{r perturb_pathifier, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('pathifier')
```

## PARADIGM

*not implemented*

```{r perturb_paradigm, echo=FALSE, fig.height=3, fig.width=8}
#util$perturb_score_plots('paradigm')
```

# ROC curves for Recovery of Perturbations

```{r perturb_all, echo=FALSE}
roc$do_plot()
```

# Stability of Scores in Basal Expression

```{r stability, echo=FALSE}
stability$do_plot('gdsc_matrix.RData')
```

# Correlation of Pathway Scores in Basal Expression

## Signaling Footprints

```{r cor_speed_matrix, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('speed_matrix')
```

## Gene Ontology

```{r cor_go, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('gsea_go')
```

## Reactome

```{r cor_reactome, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('gsea_reactome')
```

## BioCarta

```{r cor_speed_biocarta, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('gsea_biocarta')
```

## SPIA

```{r cor_spia, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('spia')
```

## Pathifier

*some error*

```{r cor_pathifier, echo=FALSE, fig.height=3, fig.width=8}
#util$cor_plots('pathifier')
```

## PARADIGM

```{r cor_paradigm, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('paradigm')
```


# Discussion

... write some thoughts here ...
