# Pathway scores in Basal and Perturbed Condition

```{r setup, include=FALSE}
library(dplyr)
b = import('base')
io = import('io')
ar = import('array')
st = import('stats')
plt = import('plot')
util = import('./util_1')

overview = import('../analyses/speed_overview/plot_bars')
cluster = import('../analyses/speed_raw/plot')
roc = import('../analyses/speed_raw/roc')
stability = import('../analyses/speed_bootstrap/stability_plot')
```

## Size of the Data Set

```{r dataset_size, echo=FALSE}
overview$do_plot()
```

## Global Structure of Perturbation Experiments

```{r global_structure, echo=FALSE}
cluster$do_plot()
```

## Statistical significance of signature genes

```{r sig_sign, echo=FALSE}
#TODO: add the worst (mean?) FDR for the signature genes
# like this, we can put a lower bound on the significance
```

## Pathway scores of different methods assigned to perturbations

Below are comparisons of the associations between input experiments and a
higher pathway score (left), as well as the scores assigned to each individual
perturbation experiment in our database (right).

We calculate linear associations using the *lm* function in R between a
pathway coefficients matrix (of the shape experiments x pathways, with the
value 1 where a pathway was activated, -1 where a pathway was
inhibited, and 0 otherwise; here, no cross-talk was modelled
explicitly) and the pathway scores obtained by using the different
methods. We then plot the result of the associations as a
heatmap where the perturbed pathway and the obtained pathway score are on the
axes, and the colour (blue for positive and red for negative correlation)
indicates the Wald statistic between them. We annotate combinations where
the p-value is below 10^-5^ with `*` and where it is 10^-10^ with `***`.

As a method for inspection of the pathway scores of the individual
experiments, We plot a heatmap using the R package *pheatmap*
with the pathways scores in rows, experiments in columns, and the
colour indicates the relative activation (scores scaled within experiments
across pathways) of each
pathway. In order to make activations and inhibitions more
comparable, we take the negative pathway score for each inhibition
and annotate the columns separately indicating which pathway was
perturbed, and which kind of perturbation it was and we ordered the pathway in
the following order: EGFR, H2O2, Hypoxia, JAK-STAT, MAPK, NFkB, PI3K,
TGFb, TNFa, Trail, VEGF, p53).



### Signaling Footprints

```{r perturb_speed_matrix, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('speed_matrix')
```

### Gene Ontology

```{r perturb_go, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('gsea_go')
```

### Reactome

```{r perturb_reactome, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('gsea_reactome')
```

### BioCarta

```{r perturb_biocarta, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('gsea_biocarta')
```

### SPIA

```{r perturb_spia, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('spia')
```

### Pathifier

```{r perturb_pathifier, echo=FALSE, fig.height=3, fig.width=8}
util$perturb_score_plots('pathifier')
```

### PARADIGM

*not implemented*

```{r perturb_paradigm, echo=FALSE, fig.height=3, fig.width=8}
#util$perturb_score_plots('paradigm')
```

## Consensus signature corresponds more closely to perturbations than pathway methods

consensus not perfect, GEX sig confounded, will not be 100% with simple model (only when overfitted)

```{r perturb_all, echo=FALSE}
roc$do_plot()
```

## Stability of Scores in Basal Expression

```{r stability, echo=FALSE}
stability$do_plot('gdsc_matrix.RData')
```

## Correlation of Pathway Scores in Basal Expression

Using basal gene expression, we calculated the
Pearson correlation across all samples and tissues, and plotted the correlation
matrix between each pathway combination using the R package *corplot* for
GDSC and TCGA data separately.

Below are the correlation for TCGA primary tumour data (left) and GDSC cell lines (right) for
perturbation-response genes in basal expression. Sign indicated by colour of
the points (positive correlation blue, negative correlation red), strength by
colour shade and size of the point.

### Signaling Footprints

```{r cor_speed_matrix, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('speed_matrix')
```

### Gene Ontology

```{r cor_go, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('gsea_go')
```

### Reactome

```{r cor_reactome, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('gsea_reactome')
```

### BioCarta

```{r cor_speed_biocarta, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('gsea_biocarta')
```

### SPIA

```{r cor_spia, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('spia')
```

### Pathifier

```{r cor_pathifier, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('pathifier')
```

### PARADIGM

```{r cor_paradigm, echo=FALSE, fig.height=3, fig.width=8}
util$cor_plots('paradigm')
```


## Discussion

... write some thoughts here ...
